---
title: "Using PrismJS with {xaringan}"
author: "Adam Shen"
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    css: css/xaringan-themer.css
    mathjax: null
    nature:
      highlightStyle: null
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)

htmltools::tagList(
  htmltools::htmlDependency(
    name = "prismr",
    version = "1.28.0",
    src = "libs/prismr",
    script = "prismr.js",
    stylesheet = "adam-one-dark.css"
  )
)

old_hooks <- fansi::set_knit_hooks(
  knitr::knit_hooks,
  which = c("output", "message", "warning", "error")
)


options(crayon.enabled=TRUE)
```

count: false
class: title-slide, center, middle

# Using PrismJS with {`xaringan`} via `prismr`

.pull-left[

`r fontawesome::fa("code-branch")` [`prismr` (Github repo)](github.com/adamoshen/prismr)

]

.pull-right[

`r fontawesome::fa("github")` [@adamoshen](github.com/adamoshen)

`r fontawesome::fa("twitter")` [@adamoshen](twitter.com/adamoshen)

]

---

class: language-r

# PrismJS works with `fansi` knit hooks

```{r, message=TRUE}
library(tidyverse)
```

---

class: language-r

# PrismJS works with `fansi` knit hooks

```{r}
tidyverse_logo()
```

---

class: language-r

# Example code and output

```{r}
new_penguins <- palmerpenguins::penguins %>%
  drop_na() %>%
  mutate(bill_length_mean1 = mean(bill_length_mm)) %>%
  group_by(species, island) %>%
  mutate(bill_length_mean2 = mean(bill_length_mm)) %>%
  ungroup() %>%
  select(species, island, contains("bill_length"))

new_penguins
```

---

# Highlighting code inline without defining a slide class

You can also highlight inline code using the usual dot-classes, with the class being "language-r".

So I thought I'd use .language-r[`dplyr::group_by()`] followed by .language-r[`dplyr::mutate()`].
Of course, afterward I would also .language-r[`dplyr::ungroup()`].

You can also use the class "language-none" to disable highlighting of inline code.

So I thought I'd use .language-none[`dplyr::group_by()`] followed by .language-none[`dplyr::mutate()`].
Of course, afterward I would also .language-none[`dplyr::ungroup()`].

---

# You may be interested in using the `layout` feature

Recall that you can define slide classes for all subsequent slides using `layout`. As usual, you
can also combine classes together.

```
class: center, inverse, language-r
layout: true
```

Read more about slide layouts
[`r fontawesome::fa("external-link-alt")`](https://github.com/yihui/xaringan/wiki/Slide-layouts).

---

class: language-r

# Limitations

One known limitation is that adding comments will highlight the remainder of the chunk as a comment
[`r fontawesome::fa("external-link-alt")`](https://github.com/adamoshen/prismr/issues/2).

```{r, eval=FALSE}
1 + 2

# Everything past this point will be highlighted as a comment
mean(c(1, 4, 5, 6))
data(starwars, package="dplyr")
```

Solutions:

- Only include comments on the last line, or adjacent to the last line of code
- Don't include comments in your code chunk
- Start a new code chunk after commenting
- Use plain text on the slide to explain your code
- Line highlighting still works though (see next slide)

---

class: language-r

# This is something complicated

.pull-left[

```{r, eval=FALSE}
ggplot(
  mpg, #<<
  
  aes(x = displ, y = hwy, colour = class)
) +
  geom_point()
```

]

.pull-right[

Construct a plot using the .language-none[`mpg`] data set.

]

--

.pull-left[

```{r, eval=FALSE}
ggplot(
  mpg,
  
  aes(x = displ, y = hwy, colour = class) #<<
) +
  geom_point()
```

]

.pull-right[

The x-axis will represent engine displacement.

The y-axis will represent highway miles per gallon.

Geoms will be coloured by type of vehicle

]

--

.pull-left[

```{r, eval=FALSE}
ggplot(
  mpg,
  
  aes(x = displ, y = hwy, colour = class)
) +
  geom_point() #<<
```

]

.pull-right[

Draw points on the canvas.

]
